sensi_an_rotations1. Start of sensitivity analysis for rotations.

### 2018-03-20 inst/documents/rotations_notes.Rmd
thorough check

@@

@@

~ how to deal with 'overshooting' leading to more generations under 50 when rotation intervals are shorter ? 
run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, add_gens_under50=TRUE)
run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, add_gens_under50=TRUE)

@@

~ something not right, when run with coverage=0 still get increase in resistance in treated areas, even though technically there is no treated area, maybe I just need to force coverage to be > 0 :
run_rot( n_insecticides = 3 , max_gen = 150 , start_freqs = 0.01 , rot_interval = 0 , eff = 0.8 , dom_sel = 0.5 , dom_cos = 0.5 , rr = 0.5 , expo_hi = 0.5 , coverage = 0 , migration = 0.01 , cost = 0.01 , logy = TRUE no_r_below_start = TRUE min_rwr_interval = 5 )

>>>>

*19/3/2018* on train to liverpool

trying to understand overshooting issue

but doesen't look like much time over the 0.5 thresh for the rotations can explain why they take longer. The short rotations only get to 0.5 towards the end.

Can I refine this example to understand what may be happening ?

turn off log ?

dfseq <- run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
dfrot <- run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE, logy=FALSE)

2 insecticides
dfseq <- run_rot(rot_interval=0, max_gen = 70, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, n_insecticides=2)
dfrot <- run_rot(rot_interval=6, max_gen = 70, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, n_insecticides=2)
rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE, logy=FALSE)
gens_under_thresh(dfrot)
gens_under_thresh(dfseq)

This does go over the 0.5 threshold ? can I count this example in sensi_an_rotations ?

gens_under_thresh(dfrot) : 49
gens_under_thresh(dfseq) : 46

When there is no cost they are identical (39) :
dfseq <- run_rot(rot_interval=0, max_gen = 70, migration=0.01, expo_hi=0.8, eff=0.8, cost=0, coverage=1, n_insecticides=2)
dfrot <- run_rot(rot_interval=6, max_gen = 70, migration=0.01, expo_hi=0.8, eff=0.8, cost=0, coverage=1, n_insecticides=2)
rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE, logy=FALSE)
gens_under_thresh(dfrot)
gens_under_thresh(dfseq)

but they still look diff in plots, i think because stoppping criteria are different in plots for seq & rot.

# weirdly when I went to 3 insecticides, & cost 0.1
# gen calc gives same result for each at 69 and plots finish at same point ?
# aha! that was because max_gen=70 so max result could be 69 i.e. no insectidides in either scenario reached threshold

dfseq <- run_rot(rot_interval=0, max_gen = 100, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, n_insecticides=3)
dfrot <- run_rot(rot_interval=6, max_gen = 100, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1, n_insecticides=3)
rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE, logy=FALSE)
gens_under_thresh(dfrot)
gens_under_thresh(dfseq)

1 20-20 = 0
2 insecticides 46 - 49 = 6
3 insecticides difference is 83-79 = 4
4 : 147 - 124 = 23
5 : 218 - 200 = 18
6 : 437 - 391 = 46

? no obvious pattern

done ~ write function to count gens under thresh
done ~ add thresh line to plot
done ~ add warning to run_rot() for if threshold not reached in time

exit_rot seems not quite to be doing what I expect. e.g. if I set to just 1 insecticide sequence keeps going longer than rot (although this is a somewhat weird example)

Aha! it is min_rwr_interval that did this. It stops a change early in a rotation but this can stop the simulation from ending.

(however it shouldn't effect the gens_under_threshold)

Try seeing if the gens_under_thresh are the same when min_rwr_interval is disabled.

(also note that min_rwr_interval isn't really sposed to apply to rotations, partly a result of the structure )

setting min_rwr_interval to 1 now results if diff of 86-77 = 9 for the 3 insecticides ??

*20/3/18*
Go back to the scenarios that was very different in sensi_an_rotations_test num79

dfseq <- run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
dfrot <- run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE, logy=FALSE)
gens_under_thresh(dfrot)
gens_under_thresh(dfseq)

To enable side by side comparison

library(gridExtra)
ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE)
ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE)
grid.arrange(ggrot, ggseq, ncol = 1)

I think maybe need to go back to two insecticides to try to find small difference :

ni <- 2
cost <- 0.1
gens <- 70 
dfseq <- run_rot(rot_interval=0, max_gen=gens, migration=0.01, expo_hi=0.8, eff=0.8, cost=cost, coverage=1, n_insecticides=ni)
dfrot <- run_rot(rot_interval=6, max_gen=gens, migration=0.01, expo_hi=0.8, eff=0.8, cost=cost, coverage=1, n_insecticides=ni)
ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE)
ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE)
grid.arrange(ggrot, ggseq, ncol = 1)
gens_under_thresh(dfrot)
gens_under_thresh(dfseq)

rot:48, seq:46

Note that in the above plot other strange issues, which may be plotting probs.
~ insecticide use boxes don't join up, seem to be gaps
~ resistance keeps increasing after insecticide box has finished
~ first rotation seems shorter than later, because insecticide is classed as NA in firs tstep, is that right ?

Difference is for i2 which has 4 rots of 6, 24, only reaches threshold at end of final rot
In seq, 12 last for 2 generations less.

Trying too look at plot in more detail
dfrot2 <- run_rot(rot_interval=6, max_gen=20, migration=0.01, expo_hi=0.8, eff=0.8, cost=cost, coverage=1, n_insecticides=ni)

Not sure if I'm going to be able to fix the ribbons joining ? can I +1 to the x upper limit ?

  to_plot_in_use <- 1

  dfrot2 <- dfrot2 %>%
    mutate( i_in_use = ifelse(resist_gene==paste0('insecticide',insecticide), to_plot_in_use, NA))

  gg <- ggplot( dfrot2, aes_string(x='generation',y='resistance')) +
    geom_line( alpha=0.5, lwd=1, colour='red3' ) +
    facet_wrap('resist_gene', ncol=1) +
    geom_ribbon( aes_string(x='generation', ymin=0, ymax='i_in_use'))

  plot(gg)

I think ribbons not meeting and curve increasing after are just plotting issues that I don't know how to fix.

If a rotation is 10 generations, the first finishes at 10 and the next is shown as starting at 11. The effect of the first continues until 11. Setting i_in_use to 1 for the following generation would fix the plotting issue.

Do I need to try to add a generation 0 to store starting conditions ? NO

Generations
The conditions within a generation influence resistance which is stored in the following generation.
Therefore it's good to set insecticide for time 1, and this is consistent with later generations.

So I've fixed saving of insecticide for t1

ni <- 2
cost <- 0.1
gens <- 70 
dfseq <- run_rot(rot_interval=0, max_gen=gens, migration=0.01, expo_hi=0.8, eff=0.8, cost=cost, coverage=1, n_insecticides=ni)
dfrot <- run_rot(rot_interval=6, max_gen=gens, migration=0.01, expo_hi=0.8, eff=0.8, cost=cost, coverage=1, n_insecticides=ni)
ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE)
ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE)
grid.arrange(ggrot, ggseq, ncol = 1)
gens_under_thresh(dfrot)
gens_under_thresh(dfseq)

Still have difference between 24 & 22 for i2 in above example

# add in_use columns  
  dfrot <- dfrot %>%
    mutate( i_in_use = ifelse(resist_gene==paste0('insecticide',insecticide), 1, NA)) 
  dfseq <- dfseq %>%
    mutate( i_in_use = ifelse(resist_gene==paste0('insecticide',insecticide), 1, NA)) 
#remove NA refuge rows    
dfrot <- dfrot[dfrot$active_or_refuge=='active',]
dfseq <- dfseq[dfseq$active_or_refuge=='active',]

write_csv(dfrot,'rot.csv')
write_csv(dfseq,'seq.csv')

see : rot_seq_compare_march2018.xlsx

With cost it seems that rotation may be able to last longer than a sequence.

Could this be because decline in resistance freq is more rapid at higher frequencies.

Both increases and decreases in resistance frequency are higher at higher resistance frequencies.

But does this explain ?

Is it something to do with more frequent changes ?

Could it be because :
~ frequency changes are bigger as frequencies get higher, 
~ allowing in a sequence to reach high frequencies straight away makes it slightly quicker than see-sawing at lower frequencies
~ is their some maths to prove this ?
~ ask Ian
~ can I print an example ?

Ian said that at higher dominance the difference in change between low & high frequencies will be greater, so might expect the difference between rot and seq to be greater. Try it and see.

linputs <- list()

linputs$n_insecticides <- 2
linputs$coverage <- 1
#linputs$migration <- 0.1
linputs$cost <-     0.1 
linputs$max_gen <-  200
linputs$expo_hi <-  0.6
linputs$eff <-      0.6
linputs$dom_sel <-  1
linputs$dom_cos <-  1

linputs$rot_interval=0
dfseq <- do.call(run_rot, linputs)
linputs$rot_interval=6
dfrot <- do.call(run_rot, linputs)

ggrot <- rot_plot_resistance(dfrot, plot_refuge=FALSE, logy=FALSE)
ggseq <- rot_plot_resistance(dfseq, plot_refuge=FALSE, logy=FALSE)
library(gridExtra)
grid.arrange(ggrot, ggseq, ncol=1)
gens_under_thresh(dfrot)
gens_under_thresh(dfseq)

with higher dominance there is only 1 generation difference (31:30) may be partly because it is faster

when I reduced exposure and effectiveness then a largish difference between rot & seq did return.

@@

~ compare dfrot & dfseq for above example

~ seems that min_rwr_interval explains the plots looking different but not the measured gens_under_thresh

~ maybe come back to i_in_use ribbons not meeting issue

~ rename rot_criterion to rot_threshold

~ add in some unit-tests ?

@@

~ how to deal with 'overshooting' leading to more generations under 50 when rotation intervals are shorter ? 
dfseq <- run_rot(rot_interval=0, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
dfrot <- run_rot(rot_interval=6, max_gen = 300, migration=0.01, expo_hi=0.8, eff=0.8, cost=0.1, coverage=1)
rot_plot_resistance(dfrot, df_resanother=dfseq, plot_refuge=FALSE)

@@

### 2018-03-20 inst/documents/sensi_an_rotations1.Rmd
thorough check

    #print scenario num and inputs to doc for exploration
    cat("scenario",i,
        " expo_hi",signif(linputs$expo_hi,2),
        " eff",signif(linputs$eff,2),
        " rot_interval",linputs$rot_interval,
        "\n")

>>>>

    #print scenario num and inputs to doc for exploration
    cat("scenario",i,
        " expo_hi",signif(linputs$expo_hi,2),
        " eff",signif(linputs$eff,2),
        " rot_interval",linputs$rot_interval,
        " cost", linputs$cost,
        "\n")
